\section{References}
 
 \noindent Amari, S., Cichocki, A. and H. Yang (1996). A new learning algorithm for blind signal separation. \emph{NIPS96}. \medskip 
 
 \noindent Bell, A. J.  and  T.J. Sejnowski (1995). An information theoretic approach to blind separation and blind deconvolution. \emph{Neural Computation}, 7(6), 1129-59.\medskip
 
 \noindent Bishop, C.M. (1992). \emph{Pattern Recognition and Machine Learning}. \medskip
 
 \noindent Brillinger, D. (1982).  non Gaussian regressors. \medskip 
 
\noindent  Carvalho, C.M., N.G. Polson and J.G. Scott (2010).  The horseshoe estimator of sparse signals. \emph{Biometrika}, .  \medskip

\noindent  Carvalho, C.M., N.G. Polson and J.G. Scott (2008). Handling Sparsity via the Horseshoe.\medskip

\noindent Cemgil, A. T., C. Fevotte and S. J. Godsill (2008). Variational  and Stochastic Inference for Blind Source Separation.\medskip 

\noindent Comon, P., C. Jutten and J. Herault (1991). Blind separation of sources.   \medskip 
 
 \noindent Deco. \medskip
 
 \noindent Detommaso et al.(2019). \medskip
 
\noindent Dinh, L., D. Krueger, Y. Bengio (2014).  NICE: Non-linear Independent Components Estimation. \medskip 

\noindent Donnat, C.,  L. Tozzi and  S. Holmes (2019). Constrained Bayesian ICA for Brain Connectome inference.\medskip

\noindent Fevotte, C. and S. J. Godsill  (2004). Bayesian Approach for Blind Separation of Underdetermined Mixtures of Sparse Sources.
\emph{ICA and BSS}, 398-405.\medskip


\noindent Fevotte, C.S. J. Godsill  and P. Wolfe (2008).  Blind Separation of Sparse Signals using Jeffrey's Inverse Prior and the EM algorithm. 
\emph{ICA}, 593-600.\medskip 

\noindent Hyvarinen, A. and H. Morioka (2019). Nonlinear ICA. \medskip

\noindent Hyvarinen, A. and P. Pajunen (1999). Nonlinear Independent Component Analysis: existence and uniqueness results. \medskip

\noindent  MacKay, D.J.C.  (1992).  Bayesian Interpolation. \emph{Neural Computation}, 4(3), 415-447.\medskip

\noindent  MacKay, D.J.C.  (1994). Maximum Likelihood and Covariant Algorithms for Independent Component Analysis. \medskip

\noindent Ono Miyake \medskip

\noindent Palmer, J. (2013). Super-Gaussian \medskip 

\noindent  Pearlmutter and Parra (1996). A context-sensitive generalisation of ICA.  \medskip

\noindent  Polson, N.G. and J.G. Scott  (2012).  Non-Gaussian data augmentation \medskip

\noindent  Polson, N.G.,  J.G. Scott  and J. Windle (2013). Bayesian Inference for logistic models using P\'olya-Gamma latent variables. 
\emph{J. Amer. Stat. Assoc.},  1339-1349. \medskip

\noindent Polson, N.G. and V. O. Sokolov (2017). Bayesian Deep Learning. \emph{Bayesian Analysis}, .\medskip 

\noindent Polson, N.G. and V. O. Sokolov (2021). Deep Learning Partial Least Squares. \emph{arXiv}, .\medskip 

\noindent Valpola \medskip

\noindent Z and  Pearlmutter \medskip